---
title: "explAnalysis"
author: "Ana Francesca Stama"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(zoo)
library(fda)
```

## Učitavanje i pregled podataka

```{r}
INPUT_FILE <- "Zagreb_dnevniQ-1960-2019.csv"
INPUT_FILE_BASEFLOW <- "Zagreb_baseflow.csv"
```
```{r}
df <- read_csv(INPUT_FILE)
spec(df)
```
```{r}
df_base <- read_csv(INPUT_FILE_BASEFLOW)
spec(df_base)
df_base[, c(4, 6)]
```

```{r}
names(df) <- c("date", "flow")

df$date <- dmy(df$date)
df$year <- year(df$date)
df$baseflow <- df_base$baseflow

df[month(df$date) >= 10, ]$year <- df[month(df$date) >= 10, ]$year + 1

summary(df)
```
```{r}
df$floworig <- df$flow
df$flow <- df$flow - df$baseflow

df <- filter(df, !is.na(flow))

summary(df)
```
```{r}
df %>% group_by(year) %>% summarise(flow = max(flow), flowDesc = "peak") -> dfMax
dfMax
df %>% full_join(dfMax) -> df
```
```{r}
df
```

```{r}
#df %>% ggplot(aes(x = date, y = floworig)) + geom_line()
df %>% ggplot(aes(x = date, y = flow)) + geom_line()
```
## Problem Nove Godine

```{r}
LEFT_CUTOFF <- dmy("1.1.1960")
RIGHT_CUTOFF <- dmy("1.1.1972")

LEFT_CUTOFF
df$dekada <- year(df$date) %>% `%%`(100) %>% `%/%`(10)
unique(df$dekada)

df$godina <- year(df$date) %>% `%%`(10)
unique(df$godina)

df %>% filter(date >= LEFT_CUTOFF, date <= RIGHT_CUTOFF) %>% ggplot(aes(date, flow)) + geom_line() + 
  geom_vline(aes(xintercept = datum), data = data.frame(datum = seq(dmy("1.1.1960"), dmy("1.1.2020"), by = "year")),  col = "blue")  +
  geom_vline(aes(xintercept = date), data = filter(df, !is.na(flowDesc)),  col = "red", linetype = 2)
```


## Traženje peak-ova

```{r}
peakDates <- df %>% filter(flowDesc == "peak") %>% pull(date) %>% as.character
##peakDates <- (df %>% filter(flowDesc == "peak"))$date
peakDates
```
## Postavljanje segmenata

```{r}
df$peak_number <- as.numeric(NA)
peak_number <- 1
curr_year <- numeric()

overlap_counter <- 0

for (d in peakDates) {
 curr_year <- year(as.Date(d))
 df[df$date == as.Date(d), ]$peak_number <- peak_number
 
 # šetnja po datumima prije peak-a
 for (dPast in as.character(seq(as.Date(d) - 1, min(df$date), -1))) {
   if (!is.na(df[df$date == as.Date(dPast), ]$flowDesc)) overlap_counter <- overlap_counter + 1
   df[df$date == as.Date(dPast), ]$peak_number <- peak_number
   
   if (curr_year > year(as.Date(dPast))) {
     df[df$date == as.Date(dPast), ]$year <- curr_year
   }

   if (round(df[df$date == as.Date(dPast), ]$floworig, 0) ==round(df[df$date == as.Date(dPast), ]$baseflow, 0)) {
     df[df$date == as.Date(dPast),]$flowDesc <- "waveStart"
     break
   } else df[df$date == as.Date(dPast), ]$flowDesc = "leftSegment"
 }
  
 
  # šetnja po datumima poslije peak-a
  for (dFuture in as.character(seq(as.Date(d) + 1, max(df$date), 1))) {
    if (!is.na(df[df$date == as.Date(dFuture), ]$flowDesc)) overlap_counter <- overlap_counter + 1
    
   df[df$date == as.Date(dFuture), ]$peak_number <- peak_number
   
   if (curr_year < year(as.Date(dFuture))) {
     df[df$date == as.Date(dFuture), ]$year <- curr_year
   }
   
   if (round(df[df$date == as.Date(dFuture), ]$floworig, 0) == round(df[df$date == as.Date(dFuture), ]$baseflow, 0)) {
     df[df$date == as.Date(dFuture),]$flowDesc <- "waveEnd"
     break
   } else df[df$date == as.Date(dFuture), ]$flowDesc = "rightSegment"
  }
 
 curr_year <- curr_year + 1
 peak_number <- peak_number + 1
}


startDates <- df %>% filter(flowDesc == "waveStart") %>% pull(date) 
endDates <- df %>% filter(flowDesc == "waveEnd") %>% pull(date) 

df$flowDay <- as.numeric(NA)

for (i in 1:length(startDates)) {
  df[df$date >= startDates[i] & df$date <= endDates[i], "flowDay"] <- 
    0:(as.numeric(endDates[i] - startDates[i]))
}
df
```
```{r}
df %>% filter(flowDesc == "peak") %>% dplyr::select(year, peakDay = flowDay) %>% 
   full_join(df) -> df
df
```
```{r}
df %>% mutate(flowDayShifted = flowDay - peakDay) -> df
df
```


## Checkpoint
```{r}
df_checkpoint_1 <- df
```

## Fix zbog preklapanja
```{r}
df <- df_checkpoint_1
```
```{r}
# nisam sigurna čemu ovo
leftmostDay <- df %>% filter(flowDesc=="waveStart") %>% pull(flowDayShifted) %>% min
rightmostDay <- df %>% filter(flowDesc=="waveEnd") %>% pull(flowDayShifted) %>% max
leftmostDay
rightmostDay
```
```{r}
waveStartDates <- df %>% filter(flowDesc == "waveStart") %>% pull(date) %>% as.character
waveEndDates <- df %>% filter(flowDesc == "waveEnd") %>% pull(date) %>% as.character

for (d in waveStartDates) {
   dd <- as.Date(d)
   yearNo <- df %>% filter(date == d) %>% pull(year)
   peakNo <- df %>% filter(date == d) %>% pull(peak_number)
   leftDay <- df %>% filter(date == d) %>% pull(flowDayShifted)
   #print(str_c(year, peakNo, leftDay, sep = " "))
   
   while (leftDay > leftmostDay) {
      leftDay <- leftDay - 1
      dd <- dd - 1
      df %>% mutate(is.date = date == dd,
                    flowDesc = ifelse(is.date, "extensionLeft", flowDesc), 
                    peak_number = ifelse(is.date, peakNo, peak_number),
                    year = ifelse(is.date, yearNo, year),
                    flowDayShifted = ifelse(is.date, leftDay, flowDayShifted),
                    flow = ifelse(is.date, 0, flow)) -> df
   } 
}


for (d in waveEndDates) {
   dd <- as.Date(d)
   yearNo <- df %>% filter(date == d) %>% pull(year)
   peakNo <- df %>% filter(date == d) %>% pull(peak_number)
   rightDay <- df %>% filter(date == d) %>% pull(flowDayShifted)
   
   while (rightDay < rightmostDay) {
      rightDay <- rightDay + 1
      dd <- dd + 1
      df %>% mutate(is.date = date == dd,
                    flowDesc = ifelse(is.date, "extensionRight", flowDesc), 
                    peak_number = ifelse(is.date, peakNo, peak_number),
                    year = ifelse(is.date, yearNo, year),
                    flowDayShifted = ifelse(is.date, rightDay, flowDayShifted),
                    flow = ifelse(is.date, 0, flow)) -> df
   } 
}

df$is.date <- NULL
df
```

## Checkpoint
```{r}
df_checkpoint_2 <- df
```

```{r}
df <- filter(df, !is.na(flowDayShifted))
df
```

```{r}
df %>% group_by(year) %>% summarise(min = min(flowDayShifted), max = max(flowDayShifted))
```
```{r}
df %>% pull(year) %>% unique %>% length
```

## Checkpoint
```{r}
df_checkpoint_3 <- df
df
```



## Eksploratorna analiza
```{r}
#svi valovi na jednom grafu
df %>% ggplot(aes(x = flowDayShifted, y = flow, col = as.factor(peak_number))) + geom_line() +
  theme(legend.position = "none")
```
```{r}
df %>% ggplot(aes(x = flowDayShifted, y = flow)) + geom_line() + facet_wrap(~year,  ncol=10)
```


## Normalizacija y vrijednosti

```{r}
# "area under curve"
AUC <- function(x, y) {
   sum(diff(x)*rollmean(y,2))
}
```
```{r}
df %>% group_by(year) %>% summarise(auc = AUC(flowDayShifted, flow)) %>% full_join(df) -> df
df$flowScaled <- df$flow / df$auc

df %>% ggplot(aes(x = flowDayShifted, y = flowScaled)) + geom_line() + facet_wrap(~year,  ncol=10)
```

## B-splines

```{r}
POINTS <- 55

Bspline.basis.4 <- create.bspline.basis(c(0, POINTS), 4)
Bspline.basis.10 <- create.bspline.basis(c(0, POINTS), 10)
fourier.basis.4 <- create.fourier.basis(c(0, POINTS), 4)
fourier.basis.10 <- create.fourier.basis(c(0, POINTS), 10)

plot(Bspline.basis.4)
plot(Bspline.basis.10)
plot(fourier.basis.4)
plot(fourier.basis.10)
```
```{r}
# sanity check

obs <- df %>% filter(year == 1960) %>% pull(flowScaled)
#obs
flow.fd.1960.b4 <- smooth.basis(y = obs, fdParobj=Bspline.basis.4)
flow.fd.1960.b10 <- smooth.basis(y = obs, fdParobj=Bspline.basis.10)
flow.fd.1960.f4 <- smooth.basis(y = obs, fdParobj=fourier.basis.4)
flow.fd.1960.f10 <- smooth.basis(y = obs, fdParobj=fourier.basis.10)
plot(obs)
lines(flow.fd.1960.b4, col = "red")
lines(flow.fd.1960.b10, col = "blue")
lines(flow.fd.1960.f4, col = "purple")
lines(flow.fd.1960.f10, col = "green")
```
```{r}
# sanity check 2

obs <- df %>% filter(year == 1964) %>% pull(flowScaled)
flow.fd.1960.b4 <- smooth.basis(y = obs, fdParobj=Bspline.basis.4)
flow.fd.1960.b10 <- smooth.basis(y = obs, fdParobj=Bspline.basis.10)
flow.fd.1960.f4 <- smooth.basis(y = obs, fdParobj=fourier.basis.4)
flow.fd.1960.f10 <- smooth.basis(y = obs, fdParobj=fourier.basis.10)
plot(obs)
lines(flow.fd.1960.b4, col = "red")
lines(flow.fd.1960.b10, col = "blue")
lines(flow.fd.1960.f4, col = "purple")
lines(flow.fd.1960.f10, col = "green")
```
```{r}
obsAll <- df %>% pull(flowScaled) %>% matrix(ncol = POINTS, byrow = T)


flow.fd <- smooth.basis(y = t(obsAll), fdParobj=Bspline.basis.4)
plot(flow.fd)
flow.fd <- smooth.basis(y = t(obsAll), fdParobj=Bspline.basis.10)
plot(flow.fd)
obsAll <- df %>% pull(flowScaled) %>% matrix(ncol = POINTS, byrow = T)
flow.fd <- smooth.basis(y = t(obsAll), fdParobj=fourier.basis.4)
plot(flow.fd)
obsAll <- df %>% pull(flowScaled) %>% matrix(ncol = POINTS, byrow = T)
flow.fd <- smooth.basis(y = t(obsAll), fdParobj=fourier.basis.10)
plot(flow.fd)
```

## K-means clustering

```{r}
wavesCoef.bspline <- as_tibble(t(coef(flow.fd)))
wavesCoef.fourier <- as_tibble(t(coef(flow.fd)))

k3 <- kmeans(wavesCoef.fourier, centers = 3, nstart = 25)
clustersFourier <- k3$cluster

k3 <- kmeans(wavesCoef.bspline, centers = 3, nstart = 25)
clustersBspline <- k3$cluster
```

```{r}
df %>% filter(flowDesc == "peak") %>% dplyr::select(peak_number) %>% 
   mutate(clusterFourier = as.factor(clustersFourier),
          clusterBspline = as.factor(clustersBspline)) %>% full_join(df) -> df
```
```{r}
df %>% ggplot(aes(x = flowDayShifted, y = flowScaled, col = clusterFourier)) + geom_line() + facet_wrap(~year,  ncol=10)
df %>% ggplot(aes(x = flowDayShifted, y = flow, col = clusterBspline)) + geom_line() + facet_wrap(~year,  ncol=10)
```

## Komentari
- volumen
- two-stage clustering approaches
- "The threshold for the peak discharge was chosen iteratively to fulfill a target condition of four events per year on average"
- "Flood events that ended in a secondary peak were removed
automatically from the data set to ensure that events triggered by two independent convective precipitation events were not considered as one event"


















